Automatically Splitting a Two-Stage Lambda Calculus

### 2015-10-15 paper/relatedwork.tex
stuff on values

Researchers have explored other staging transformations that, 
like splitting, partition an input two-stage program into two components,
on corresponding to each stage.  
In particular, \emph{binding time separation}\,\cite{Mogensen89a} 
(also called \emph{program bifurcation}\,\cite{DeNiel91})
has been used as a preprocessor step in partial evaluators,
allowing efficient specialization of programs with 
mixed-stage data structures without changes to the specializer itself.
Notably, the grammar-based binding-time specifications used in binding 
time separation were capable of describing
data structures with purely-static, purely-dynamic, and mixed stage
content, much like the type system of \lang\
(though this correspondence is less clear without our addition of 
\bbonep\ and $\curr$).

>>>>

Researchers have explored other staging transformations that, 
like splitting, partition an input two-stage program into two components,
on corresponding to each stage.  
In particular, \emph{binding time separation}\,\cite{Mogensen89a} 
(also called \emph{program bifurcation}\,\cite{DeNiel91})
has been used as a preprocessor step in partial evaluators,
allowing efficient specialization of programs with 
mixed-stage data structures without changes to the specializer itself.
Notably, the grammar-based binding-time specifications used in binding 
time separation are capable of describing
data structures with purely-static, purely-dynamic, and mixed-stage
content, much like the type system of \lang\
(though this correspondence is less clear without our addition of 
\bbonep\ and $\curr$).

### 2015-10-15 paper/semantics.tex
changes all over
stuff on values

The abstract syntax of \lang\ is presented as a grammar in \ref{fig:grammar}.  
Note that some production rules are parameterized by world.
Our syntax also distinguishes values from general terms, 
with the underline term constructor (``$\exv v$'') explicitly noting the inclusion of values into general terms.
There are three forms of value: \emph{partial values} for \bbonem, \emph{ground values} for \bbonep, and \emph{residuals} for \bbtwo.
All represent computations which have no remaining work in the first stage.

@@

The statics and dynamics of \lang\ are related by the type safety theorems
in \ref{fig:typeSafety}, again annotated by world.
In all cases, $\Gamma$ may be any list of variable bindings at world \bbtwo,
representing the second stage binders under which we are evaluating.
Note how the progress theorem for world \bbonem\ states that every well-typed term must take 
either a standard step or a hoisting step.

>>>>

The abstract syntax of \lang\ is presented as a grammar in \ref{fig:grammar}.  
To simplify the upcoming translation in \ref{sec:splitting}, 
we have chosen to statically distinguish between values and general computations,
using an underline constructor (``$\exv v$'') which explicitly note the parts of a computation that have been reduced to a value.
\footnote{In this presentation, the value/computation distinction is encoded intrinsically at the syntactic level.
However, we found it more convenient in the formal Twelf implementation to maintain the distinction with an extrinsic judgement.}
Moreover, the value/computation distinction interacts non-trivially with the three possible world classifications.
As a result, we end up with six classes of term in the grammar.

The key feature of all three forms of value is that they have no remaining work in the first stage.
That is, all of the first-stage portions of a value are fully reduced (except the bodies of first-stage functions).
In the case of values at \bbonep, which we call \emph{ground values}, this collapses to just the standard notion of values in a monostage language.
In the case of values at \bbtwo, which we call \emph{residuals}, this collapses to just standard monostage terms.
Lastly, values at \bbonem\ end up with more of a mixed character, so we call them \emph{partial values}.

@@

The statics and dynamics of \lang\ are related by the type safety theorems below, again annotated by world.
In all cases, $\Gamma$ may be any list of variable bindings at world \bbtwo,
representing the second stage binders under which we are evaluating.
Note how the progress theorem for world \bbonem\ states that every well-typed term must take 
either a standard step or a hoisting step.
